---
title: "Exploring Prudential Life Insurance Assessment"
author: "wittmann"
date: "November 30, 2015; last update Dec 2"
output: html_document
---

*Please let me know of any suggestions or errors*  

This is inspired by the fantastic scripts: <br>
<https://www.kaggle.com/darraghdog/springleaf-marketing-response/explore-springleaf> <br>
<https://www.kaggle.com/thie1e/rossmann-store-sales/exploratory-analysis-rossmann>


Read in the data:

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(gridExtra)

train <- read.table("../input/train.csv", sep=",", header=TRUE)
test <- read.table("../input/test.csv", sep=",", header=TRUE)

```

Separate categorical, continous and discrete variables

```{r}
cat.var.names <- c(paste("Product_Info_", c(1:3,5:7), sep=""), paste("Employment_Info_", c(2,3,5), sep=""),
                   paste("InsuredInfo_", 1:7, sep=""), paste("Insurance_History_", c(1:4,7:9), sep=""), 
                   "Family_Hist_1", paste("Medical_History_", c(2:14, 16:23, 25:31, 33:41), sep=""))
cont.var.names <- c("Product_Info_4", "Ins_Age", "Ht", "Wt", "BMI", "Employment_Info_1", "Employment_Info_4", 
                    "Employment_Info_6", "Insurance_History_5", "Family_Hist_2", "Family_Hist_3", "Family_Hist_4", 
                    "Family_Hist_5")
disc.var.names <- c("Medical_History_1", "Medical_History_15", "Medical_History_24", "Medical_History_32", 
                    paste("Medical_Keyword_", 1:48, sep=""))

train.cat <- train[, cat.var.names]
test.cat <- test[, cat.var.names]

train.cont <- train[, cont.var.names]
test.cont <- test[, cont.var.names]

train.disc <- train[, disc.var.names]
test.disc <- test[, disc.var.names]

train.cat <- as.data.frame(lapply(train.cat, factor))
test.cat <- as.data.frame(lapply(test.cat, factor))

```

Let's take a first look at the data:

```{r}
str(train.cont)
str(train.disc)
str(test.cont)
str(test.disc)

summary(train.cont)
summary(train.disc)

summary(test.cont)
summary(test.disc)

head(train)
head(test)

```

Take a look at the data with categorical features:

```{r}
str(train.cat)
str(test.cat)

summary(train.cat)
summary(test.cat)

```

Contrary to the suggested separation of the variables, it seems reasonable to use the variables Medical_History_2  and Medical_History_10 as continuous.

What are the dimensions of the datasets?

```{r}
cat("Train data has", nrow(train), "rows and", ncol(train), "columns! \n")
cat("Test data has", nrow(test), "rows and", ncol(test), "columns! \n")

```

In the above structure commands we saw missing data, how much is it?


```{r}
sum(is.na(train)) / (nrow(train) * ncol(train))
sum(is.na(test)) / (nrow(test) * ncol(test))

apply(train, 2, function(x) { sum(is.na(x)) })
apply(test, 2, function(x) { sum(is.na(x)) })

```

Can we see any different missing data structure depending on the response?

```{r}
train.na.per.response <- sapply(sort(unique(train$Response)), function(x) { apply(train[train$Response == x, ], 2, function(y) { sum(is.na(y)) }) })
train.na.per.response
round(colSums(train.na.per.response) / sum(train.na.per.response), digits=4)

```

Data with response equal 8 has the most and response equal 3 the least missing data.

Plot the missingness structure (only for those features with missing data) where the feature with
the most missing data is on top and with the least missing data on bottom.

```{r, fig.width=12, fig.height=10}
plotMissingnessStr <- function(data.in, title=NULL) {
  r <- as.data.frame(ifelse(is.na(data.in), 0, 1))
  r <- r[,order(colMeans(is.na(data.in)))]
  dat.tmp <- expand.grid(list(x=1:nrow(r), y=colnames(r)))
  dat.tmp$r <- as.vector(t(r))

  ggplot(dat.tmp) + geom_tile(aes(x=x, y=y, fill=factor(r))) + scale_fill_manual(values=c("black", "white"), name="Missing\n(0=Yes, 1=No)") + theme_light() + ylab("") + xlab("") + ggtitle(title)
}

plotMissingnessStr(data.in=train[,apply(train, 2, function(x) { sum(is.na(x)) > 0 })], title="Train data set")
plotMissingnessStr(data.in=test[,apply(test, 2, function(x) { sum(is.na(x)) > 0 })], title="Test data set")

```


Are there any duplicate rows?

```{r}
cat("Train data set - Number of duplicated rows:", nrow(train) - nrow(unique(train)), "\n")
cat("Test data set - Number of duplicated rows:", nrow(test) - nrow(unique(test)), "\n")

```

Are there any constant columns?

```{r}
train.const <- sapply(train, function(x) { length(unique(x)) == 1 })
test.const <- sapply(test, function(x) { length(unique(x)) == 1 })
cat("Train data set - Number of constant columns:", sum(train.const), "\n")
cat("Test data set - Number of constant columns:", sum(test.const), "\n")
```

Plot histograms of categorical variables

```{r}
plotHist <- function(data.in, i) {
  data <- data.frame(x=data.in[,i])
  p <- ggplot(data=data, aes(x=factor(x))) + geom_histogram() + xlab(colnames(data.in)[i]) + theme_light() + 
    theme(axis.text.x=element_text(size=8))
  return (p)
}

doPlots <- function(data.in, fun, ii, ncol=3) {
  pp <- list()
  for (i in ii) {
    p <- fun(data.in=data.in, i=i)
    pp <- c(pp, list(p))
  }
  do.call("grid.arrange", c(pp, ncol=ncol))
}

doPlots(data.in=train.cat, fun=plotHist, ii=1:4, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=5:8, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=9:12, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=13:16, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=17:20, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=21:24, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=25:28, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=29:32, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=33:36, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=37:40, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=41:44, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=45:48, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=49:52, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=53:56, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=57:60, ncol=2)
doPlots(data.in=train.cat, fun=plotHist, ii=61, ncol=2)

```

Densities of continous Features

```{r, warning=FALSE}
train.cont <- data.frame(train.cont, Response=train$Response)
plotDensity <- function(data.in, i) {
  data <- data.frame(x=data.in[,i], Response=data.in$Response)
  p <- ggplot(data) + #geom_density(aes(x=x, colour=factor(Response))) + 
    geom_line(aes(x=x), stat="density", size=1, alpha=1.0) +
    xlab(colnames(data.in)[i]) + theme_light()
  return (p)
}

doPlots(data.in=train.cont, fun=plotDensity, ii=1:4, ncol=2)
doPlots(data.in=train.cont, fun=plotDensity, ii=5:8, ncol=2)
doPlots(data.in=train.cont, fun=plotDensity, ii=9:12, ncol=2)
doPlots(data.in=train.cont, fun=plotDensity, ii=13, ncol=2)

```


Boxplots of continous Features depending on Response

```{r, warning=FALSE}
plotBox <- function(data.in, i) {
  data <- data.frame(y=data.in[,i], Response=data.in$Response)
  p <- ggplot(data, aes(x=factor(Response), y=y)) + geom_boxplot() + ylab(colnames(data.in)[i]) + theme_light()
  return (p)
}

doPlots(data.in=train.cont, fun=plotBox, ii=1:4, ncol=2)
doPlots(data.in=train.cont, fun=plotBox, ii=5:8, ncol=2)
doPlots(data.in=train.cont, fun=plotBox, ii=9:12, ncol=2)
doPlots(data.in=train.cont, fun=plotBox, ii=13, ncol=2)

```


Take a look at the response

```{r}
ggplot(train) + geom_histogram(aes(factor(Response))) + xlab("Response") + theme_light()

```


